---
title: "Homework 3"
author: "Lisa Eisler"
date: "10/8/2019"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)

knitr::opts_chunk$set(
  fig.width = 6,
  fig.asp = .6,
  out.width = "90%"
)

theme_set(theme_bw() + theme(legend.position = "bottom"))
```


## Problem 1

*Read in instacart data*

```{r}
library(p8105.datasets)
data("instacart")
instacart_data = instacart
```

The dataset from Instacart reflects online grocery orders. The data contains `r ncol(instacart_data)` variables and `r nrow(instacart_data)` entries describing individual product orders. Each entry refers to a specific product ordered, assigned a product ID and description, and further lists the time of day of the overall grocery order in which that product was included, the number of days since a customer's prior grocery order, and categorizes the product by aisle and department.

*How many aisles are there, and which aisles are most items ordered from?*

```{r}
instacart_aisle = instacart_data %>%
  select(aisle) %>%
  count(aisle, name = "n_aisle") %>%
  arrange(desc(n_aisle))
```

There are `r nrow(instacart_aisle)` aisles. Most orders are from the "Fresh Vegetables" aisle, followed in popularity by the "Fresh Fruits" and "Packaged Vegetables Fruits" aisles.

*Make a plot that shows the number of items ordered in each aisle, limiting this to aisles with more than 10000 items ordered.*

```{r}
instacart_aisle_plot = instacart_aisle %>%
  filter(n_aisle > 10000) %>%
  mutate(
    rank = rank(-n_aisle)
  )

ggplot(instacart_aisle_plot, aes(x = rank, y = n_aisle)) +
    geom_point(aes(color = aisle)) +
    labs(
      title = "Popular Aisles",
      x = "Aisle Popularity Ranking",
      y = "Number of Orders",
      caption = "Data from Instacart"
    )
```


