---
title: "Homework 3"
author: "Lisa Eisler"
date: "10/8/2019"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(ggridges)

knitr::opts_chunk$set(
  fig.width = 6,
  fig.asp = 1,
  out.width = "90%"
)

theme_set(theme_bw() + theme(legend.position = "bottom"))
```


## Problem 1

*Read in instacart data*

```{r}
library(p8105.datasets)
data("instacart")
instacart_data = instacart
```

The dataset from Instacart reflects online grocery orders. The data contains `r ncol(instacart_data)` variables and `r nrow(instacart_data)` entries describing individual product orders. Each entry refers to a specific product ordered, assigned a product ID and description, and further lists the time of day of the overall grocery order in which that product was included, the number of days since a customer's prior grocery order, and categorizes the product by aisle and department.

*How many aisles are there, and which aisles are most items ordered from?*

```{r}
instacart_aisle = instacart_data %>%
  select(aisle) %>%
  count(aisle, name = "n_aisle") %>%
  arrange(desc(n_aisle))
```

There are `r nrow(instacart_aisle)` aisles. Most orders are from the "Fresh Vegetables" aisle, followed in popularity by the "Fresh Fruits" and "Packaged Vegetables Fruits" aisles.

*Make a plot that shows the number of items ordered in each aisle, limiting this to aisles with more than 10000 items ordered.*

```{r}
instacart_aisle_plot = instacart_aisle %>%
  filter(n_aisle > 10000) %>%
  mutate(
    rank = rank(-n_aisle)
  )

ggplot(instacart_aisle_plot, aes(x = rank, y = n_aisle)) +
    geom_point(aes(color = aisle)) +
    labs(
      title = "Popular Aisles",
      x = "Aisle Popularity Ranking",
      y = "Number of Orders",
      caption = "Data from Instacart"
    )
```

*Make a table showing the three most popular items in each of the aisles “baking ingredients”, “dog food care”, and “packaged vegetables fruits”. Include the number of times each item is ordered in your table.*

```{r}
popular_items = instacart_data %>%
  select(aisle, product_name) %>%
  filter(aisle == "baking ingredients" | aisle == "dog food care" | aisle == "packaged vegetables fruits") %>%
  count(aisle, product_name, name = "n_product")
  
popular_items_table = popular_items %>%
  group_by(aisle) %>%
  mutate(
    rank = order(order(n_product, decreasing=TRUE))) %>%
  filter(rank < 4) %>%
  arrange(aisle, desc(n_product))

knitr::kable(
  popular_items_table, caption = 'Most Popular Items By Aisle.')
```

*Make a table showing the mean hour of the day at which Pink Lady Apples and Coffee Ice Cream are ordered on each day of the week.*

```{r}
apples_coffee_time = instacart_data %>% 
  select(product_name, order_dow, order_hour_of_day) %>%
  filter(product_name == "Pink Lady Apples" | product_name == "Coffee Ice Cream") %>%
  group_by(product_name, order_dow) %>%
  summarize(mean_hour = mean(order_hour_of_day)) %>%
  mutate(order_dow = case_when(
    order_dow == "0" ~ "Sunday",
    order_dow == "1" ~ "Monday",
    order_dow == "2" ~ "Tuesday",
    order_dow == "3" ~ "Wednesday",
    order_dow == "4" ~ "Thursday",
    order_dow == "5" ~ "Friday",
    order_dow == "6" ~ "Saturday")
    )

apple_coffee_table = apples_coffee_time %>% 
  pivot_wider(
    names_from = "order_dow",
    values_from = "mean_hour")

knitr::kable(
  apple_coffee_table, caption = 'Mean hour of the day for orders by day of the week.')
```


##Problem 2

*Read in BRFSS data*
  
```{r}
library(p8105.datasets)
data("brfss_smart2010")
```

*First, do some data cleaning*

```{r}
brfss_smart2010_data = brfss_smart2010 %>% 
  janitor::clean_names() %>% 
  filter(topic == "Overall Health") %>%
  filter(response == "Poor" | response == "Fair" | response == "Good" | response == "Very good" | response == "Excellent") %>%
  mutate(
    response_factor_ordered = factor(response, ordered = TRUE, levels = c("Poor", "Fair", "Good", "Very good", "Excellent"))
  )
```

*In 2002, which states were observed at 7 or more locations? What about in 2010?*

```{r}
brfss_location_2002 = brfss_smart2010_data %>%
  filter(year == 2002) %>%
  group_by(locationabbr) %>%
  summarise(
  n_locations = n_distinct(locationdesc))

brfss_location_2010 = brfss_smart2010_data %>%
  filter(year == 2010) %>%
  group_by(locationabbr) %>%
  summarise(
  n_locations = n_distinct(locationdesc))
```

In 2002, the states of CT, FL, NC, MA, NJ, & PA were observed at 7 or more locations. In 2010, the states of CO, PA, SC, OH, MA, NY, NE, WA, CA, MD, NC, TX, NJ, FL were observed at 7 or more locations.

*Construct a dataset that is limited to Excellent responses, and contains, year, state, and a variable that averages the data_value across locations within a state. Make a “spaghetti” plot of this average value over time within a state (that is, make a plot showing a line for each state across years – the geom_line geometry and group aesthetic will help).*

```{r}
excellent_responses_brfss = brfss_smart2010_data %>%
  filter(response == "Excellent") %>%
  rename(State = locationabbr) %>%
  group_by(State, year) %>%
  mutate(data_value_ave_state = mean(data_value)) %>%
  select(year, State, data_value_ave_state) %>%
  drop_na()

ggplot(excellent_responses_brfss, aes(x = year, y = data_value_ave_state, color = State)) +
    geom_smooth(se = FALSE) +
    labs(
      title = "Average Data Value Over Time",
      x = "Year",
      y = "Average Data Value",
      caption = "Data from BRFSS"
    )
```

*Make a two-panel plot showing, for the years 2006, and 2010, distribution of data_value for responses (“Poor” to “Excellent”) among locations in NY State.*

```{r}
data_value_response_brfss = brfss_smart2010_data %>%
  filter(year == 2006 | year == 2010) %>%
  filter(locationabbr == "NY")

ggplot(data_value_response_brfss, aes(x = response_factor_ordered, y = data_value)) + 
  geom_boxplot() +
  facet_grid(. ~ year) +
  labs(
      title = "Distribution of Data Value By Response - Box Plot",
      x = "Response",
      y = "Data Values",
      caption = "Data from BRFSS"
  )

ggplot(data_value_response_brfss, aes(x = response_factor_ordered, y = data_value)) + 
  geom_violin(aes(fill = response_factor_ordered), color = "blue", alpha = .5) + 
  stat_summary(fun.y = median, geom = "point", color = "blue", size = 4) +
  labs(
      title = "Distribution of Data Value By Response - Violin Plot",
      x = "Response",
      y = "Data Values",
      caption = "Data from BRFSS"
  )
```
